An Open-Source OpenSim Oculomotor Model for Kinematics and Dynamics Simulation

### 2018-07-19 doc/manuscript.tex
filip comments

\documentclass[11pt,a4paper,draft=false]{report}
\usepackage{amsmath,amsthm,amssymb}
\usepackage{bm,txfonts}       % bm: bold for Greek letters, txfonts -> varmathbb
\usepackage{breqn}              % automatically brake line in equations
\usepackage{cancel}
\usepackage{siunitx}            % include before arydshln
\usepackage{arydshln}         % matrix dash lines
\usepackage{booktabs}         % tables
\usepackage{color,soul}         % used for \hl{}
\usepackage[inline]{enumitem}   % use to set labels in the enumerate environment
\usepackage{fullpage}
\usepackage[pdftex]{graphicx}
\graphicspath{{./figures/}{./tikz/}}
\DeclareGraphicsExtensions{.pdf,.png,.jpg}
\usepackage{wrapfig}
\usepackage{caption}
\captionsetup{labelfont=bf}
\usepackage[caption=false,labelfont=bf]{subfig}
\usepackage{listings}
% \usepackage{algorithm}
% \usepackage{algorithmic}
\usepackage{url}
\usepackage[pdfpagelabels,pdfusetitle,colorlinks=true,pdfborder={0 0
  0},pdfauthor={Dimitar Stanev}]{hyperref}
\usepackage{natbib}
\bibliographystyle{IEEEtranN}

@@

\author{Constantinos Filip, Dimitar Stanev\footnote{Electrical and Computer
    Engineering Department, University of Patras, Greece; Corresponding
    author: \url{stanev@ece.upatras.gr}}, and Konstantinos Moustakas}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\begin{abstract}
  Studying human eye movements has significant implications for improving our
  understanding of the oculomotor system and treating various visuomotor
  disorders. An open-source musculoskeletal model of the human eye, that can be
  used for kinematics and dynamics analysis, is constructed based on the data
  reported in literature and made publicly available\footnote{SimTK project:
    \url{https://simtk.org/projects/eye}}. The model is implemented in
  \texttt{OpenSim}, which is an open-source framework for modeling and
  simulation of musculoskeletal systems. The calibration of the model parameters
  is based on physiological measurements of the human eye. The model
  incorporates an eye globe, orbital suspension tissues and six extraocular
  muscles. The excitation and activation patterns for a variety of targets can
  be calculated using the proposed closed-loop fixation controller that drives
  the model to perform saccadic movements in a forward dynamics manner. The
  controller minimizes the error between the desired saccadic trajectory and the
  predicted movement. Consequently, this model enables the investigation muscle
  activation patterns during static fixation and analyze the dynamics of various
  eye movements.
\end{abstract}

@@

Rapid and accurate eye movements are crucial for coordinated direction of
gaze~\cite{Lee2006}. Studying human eye movement has significant implications
for improving our understanding of the oculomotor system and treating visuomotor
disorders. Over the past decades, biomechanics simulation has provided the means
to analyze different human movements~\cite{Delp2007}. The same principles can be
used to analyze visual tasks by modeling the musculoskeletal properties of the
oculomotor system. Consequently, this model can be used to investigate muscle
activation patterns during static fixation, analyze the dynamics of various eye
movements, calculate metabolic costs and simulate eye disorders, such as
different forms of strabismus. Furthermore, it can be easily integrated with
available full body models in order to analyze the relation between the
vestibular and oculomotor systems.

Saccadic movements are a generated from a coordination of the six
\gls{em}. Clinical trials have provided a profound knowledge on the properties
of the \gls{em} and their line of action on the eye globe~\cite{Robinson1969a},
the resistive tension of the surrounding tissues~\cite{Collins1981} and the
length-tension relationship of the muscles~\cite{Iskander2018}. Various
computational models of the extraocular muscles and orbital mechanics have been
proposed, which provide insight and scientific bases for oculomotor
biomechanics, control of eye movement and binocular misalignment. These models
focus on the realism of muscle behavior and they were based on the viscoelastic
properties and physiological data \gls{em}.

The first 3D biomechanical model was developed by~\cite{Robinson1964a,
  Robinson1969}, who simplified the formulation by only considering the
elasticity of the \gls{em} ignoring their dynamics. The model incorporates
anatomically realistic muscle paths and empirical innervation-length-tension
relationships. To study the neural control of rapid saccadic movements, models
using anatomical and mechanical properties of \gls{em} have been developed by
accounting for the nonlinear muscle dynamics~\cite{Thelen2003a,
  Millard2013}. Such models, having the advantage of supporting dynamics
simulation, are used in conjunction with brain level
controllers~\cite{James2018}.

@@

The six \gls{em}, including four rectus muscles and two oblique muscles, are
controlled by the cranial nerves so as to track a visual target and to stabilize
the image of the object. The \gls{lr} and \gls{mr} muscles form an antagonistic
pair that produce horizontal eye movements. The \gls{sr} and \gls{ir} muscles
form the vertical antagonist pair, which mainly controls vertical eye movement
and also affects rotation about the horizontal plane and the line of sight
(secondary action) due to insertion positions. The \gls{so} muscle passes
through the cartilaginous trochlea attached to the orbital wall, which reflects
the \gls{so} path by 51 deg. The \gls{io} muscle originates from the orbital
wall anteroinferior to the globe center and inserts on the sclera posterior to
the globe equator. The primary actions of \gls{so} and \gls{io} cause rotation
of the globe around the visual axis and vertical movement.

The model relies on the passive pulley assumption. \autoref{tab:eye-muscle-path}
shows the positions of the origin, insertion and passive muscle pulleys for the
\gls{em}, defined in the local body coordinates of the eye globe. The data are
based on physiological measurements~\cite{Iskander2018}, with some minor
modification so as to prevent unrealistic muscle-surface penetration. Since no
position was documented for the origin of the \gls{so}, a point close to the
origins of the rectus muscles was chosen to match the fiber length in the
primary position of the \gls{so} muscle.

\begin{table}[ht]
  \centering
  \caption{Muscle path points for the six \gls{em} defined in the local frame of
    the eye globe (dimensions are given in meters).}\label{tab:eye-muscle-path}
  \begin{tabular}{@{}cccccccccc@{}}
    \toprule
    \textbf{Muscle}
    & \multicolumn{3}{c}{\textbf{Origin}}
    & \multicolumn{3}{c}{\textbf{Pulley}}
    & \multicolumn{3}{c}{\textbf{Insertion}} \\
    \midrule
    & \textit{\textbf{Ox}} & \textit{\textbf{Oy}} & \textit{\textbf{Oz}}
    & \textit{\textbf{Px}} & \textit{\textbf{Py}} & \textit{\textbf{Pz}}
    & \textit{\textbf{Ix}} & \textit{\textbf{Iy}} & \textit{\textbf{Iz}} \\
    \midrule
    \gls{lr} & -0.034 & 0.0006 & -0.013 & -0.0102 & 0.0003 & 0.012 & 0.0065 & 0 & 0.0101 \\
    \gls{mr} & -0.030 & 0.0006 & -0.017 & -0.0053 & 0.00014 & -0.0146 & 0.0088 & 0 & -0.0096 \\
    \gls{sr} & -0.0317 & 0.0036 & -0.016 & -0.0092 & 0.012 & -0.002 & 0.0076 & 0.0104 & 0 \\
    \gls{ir} & -0.0317 & -0.0024 & -0.016 & -0.0042 & -0.0128 & -0.0042 & 0.00805 & -0.0102 & 0 \\
    \gls{so} & 0.0082 & 0.0122 & -0.0152 & -0.030834 & 0.001145 & -0.01644 & 0.0044 & 0.011 & 0.0029 \\
    \gls{io} & 0.0113 & -0.0154 & -0.0111 & -0.00718 & -0.0135 & 0 & -0.008 & 0 & 0.009 \\
    \bottomrule
  \end{tabular}
\end{table}

@@

\begin{table}[ht]
  \centering
  \caption{Millard muscle parameters for the
    \gls{em}.}\label{tab:eye-muscle-parameters}
  \begin{tabular}{@{}cccccccccc@{}}
    \toprule
    \thead{Muscle \\ \quad}
    & \thead{Maximum Isometric \\ Force (N)}  % math \si not in bold
    & \thead{Optimal Fiber \\ Length (m)}
    & \thead{Tendon Slack \\ Length (m)}
    & \thead{Maximum Contraction \\ Velocity (m / s)} \\
    \midrule
    \gls{lr} & 1.4710 & 0.04898 & 0.0084 & 3.8483 \\
    \gls{mr} & 1.5740 & 0.04084 & 0.0038 & 4.6155 \\
    \gls{sr} & 1.1768 & 0.04487 & 0.0054 & 4.2009 \\
    \gls{ir} & 1.4269 & 0.04549 & 0.0048 & 4.1437 \\
    \gls{so} & 0.6031 & 0.03956 & 0.0265 & 4.7648 \\
    \gls{io} & 0.5590 & 0.04110 & 0.0015 & 3.5863 \\
    \bottomrule
  \end{tabular}
\end{table}

The \gls{fl} and \gls{fpe} characteristic curves of the \gls{em} differ
significantly from those of a skeletal muscle. As shown in
\autoref{fig:millard-curves}, we can fine-tune the curve parameters so as to fit
the experimental data available for the \gls{lr} muscle. The values for the
\gls{fl} and \gls{fpe} characteristic curves are summarized in
Tables~\ref{tab:eye-fl-curve} and~\ref{tab:eye-fpe-curve}, respectively.

\begin{table}
  \centering
  \parbox{.45\linewidth}{
    \centering
    \caption{Parameters of the \gls{fl} characteristic curve for the
      \gls{em}.}\label{tab:eye-fl-curve}
    \begin{tabular}{@{}cccccccccc@{}}
      \toprule
      \textbf{Parameter} & \textbf{Value} \\
      \midrule
      min norm active fiber length & 0.55 \\
      transition norm fiver length & 0.7 \\
      max norm active fiver length & 1.8 \\
      shallow ascending slope &  2.4 \\
      minimum value & 0.0 \\
      \bottomrule
    \end{tabular}
  }
  \quad
  \parbox{.45\linewidth}{
    \vspace{-1.5cm}
    \centering
    \caption{Parameters of the \gls{fpe} characteristic curve for the
      \gls{em}.}\label{tab:eye-fpe-curve}
    \begin{tabular}{@{}cccccccccc@{}}
      \toprule
      \textbf{Parameter} & \textbf{Value} \\
      \midrule
      strain at zero force & -0.18 \\
      strain at one norm force & 0.4 \\
      \bottomrule
    \end{tabular}
  }
\end{table}

\begin{figure}[ht]
  \centering
  \subfloat[\gls{fl} curve]{\includegraphics[width=0.5\textwidth,
    keepaspectratio]{active-force-length-curve.png}\label{fig:active-force-length-curve}}
  \subfloat[\gls{fl} curve]{\includegraphics[width=0.5\textwidth,
    keepaspectratio]{passive-force-length-curve.png}\label{fig:passive-force-length-curve}}
  \caption{The \gls{fl} and \gls{fpe} curve definition for the Millard muscle
    model as implemented in \texttt{OpenSim}.}\label{fig:millard-curves}
\end{figure}

\gls{em} have a higher fraction of fast twitch fibers and thus different
\gls{fv} behavior, due to different structures compared to skeletal
muscles. Despite that, the default Millard \gls{fv} curve was used for the six
\gls{em}, since the behavior of the selected muscle model depends heavily on the
maximum contraction velocity $v^{\text{max}}$. The maximum muscle contraction
velocity is tuned so as to match the peak velocity of saccadic eye movement
$\omega^{\text{max}} = 15.7 \si{\radian / \second}$
($900 \si{\degree / \second}$). Following this definition, the maximum muscle
contraction velocity is given in optimal fiber length per seconds and it is thus
different for each \gls{em}, as their optimal fiber length is different
($v^{\text{max}} = \omega^{\text{max}} r / l_o^M$). Furthermore, since the optic
nerve is much shorter that the average muscle nerve, activation and deactivation
delays ($\tau_a = \tau_d = 5 \si{\milli\second}$) are smaller. Finally, two
separate wrapping spheres for the rectus muscles and the oblique muscles were
created, to avoid abnormal changes on the \gls{fl} curve as the eye rotates.

@@

The passive connective tissues apply a restoring force, which brings the eye
back to the central position when the net force from the \gls{em} is zero. These
tissues include all non-muscular suspensory tissues, such as Tenon's capsule,
the optic nerve, the fat pad and the conjunctiva. The force-displacement curve
of the net elasticity can be represented as

\begin{equation}\label{equ:passive-tissue}
  \vec{f}_t = -k_p \vec{q} - k_c \vec{q}^3 - k_d \vec{\dot{q}}
\end{equation}
% 
where, $\vec{f}_t$ is the passive tissue forces,
$k_p= 2.225 \cdot 10^{-3} \si{\N \m / \radian}$,
$k_c= 34.53 \cdot 10^{-3} \si{\N \m / \radian^3}$ and
$k_v= 2 \cdot 10^{-3} \si{\N \m / (\radian / \s)}$ the constants and
$\vec{q}, \vec{\dot{q}} \inr{3}$ the rotational coordinates of the
model~\cite{Collins1981}. These forces are modeled using \texttt{OpenSim}'s
expression based coordinate force.

@@

A fixation controller that calculates the \gls{em} excitations required to
track a desired saccade was implemented. The controller actuates the model in a
closed-loop \gls{fd} manner. The parameters of the controller are the desired
horizontal and vertical fixation angles, the saccade onset and velocity, and the
gains of \gls{pd} tracking controller. A sigmoid function is used for generating
smooth saccade trajectories

\begin{equation}\label{equ:sigmoid}
  \begin{aligned}
    \theta_d(t) &= \frac{a}{2} \Big(\tanh(b (t - t_0)) + 1\Big) \\
    \dot{\theta}_d(t) &= \frac{a b}{2} \Big(1 - \tanh^2(b (t - t_0))\Big)
  \end{aligned}
\end{equation}
% 
where $\theta_d(t)$ and $\dot{\theta}_d(t)$ represent the desired orientation
and velocity at time $t$, $a$ the magnitude of the trajectory, $b$ the slope and
$t_0$ a time shift constant. Provided a fixation goal $\theta_g$, a desired
saccade velocity $\dot{\theta}_g$ and a saccade onset $t_g$, the parameters of
the sigmoid function are defined as $a = \theta_g$,
$b = 2 \dot{\theta}_g / \theta_g$ and $t_0 = t_s$. The \gls{pd} tracking
controller has the following form

\begin{equation}\label{equ:eye-pd-controller}
  \ddot{\theta}(t) = k_p (\theta_d(t) - \theta(t)) + k_d (\dot{\theta}_d(t) -
  \dot{\theta}(t))
\end{equation}
%
where $k_p$, $k_d $ are the tracking gains, and $\theta(t)$, $\dot{\theta}(t)$
the simulated response of the model.

The sign and magnitude of $\ddot{\theta}(t)$, both for the horizontal and
vertical displacements, are used to calculate the muscle excitation levels, by
assuming that each individual muscle rotates the eye globe in a particular
direction. \autoref{fig:simulated-saccade} depicts the simulated coordinates,
speeds and estimated \gls{em} excitation levels that reproduce the desired
saccade trajectory for different model parameters.

\begin{figure}[ht]
  \centering
  \includegraphics[width=.7\textwidth]{eye-model.png}
  \caption{Eye model with a fixation $\theta_H = -15 \si{\degree}$,
    $\theta_V = 15 \si{\degree}$ during simulation. Blue denotes low and red
    high muscle excitation levels.}\label{fig:model}
\end{figure}

\begin{figure}[ht]
  \vspace{-1.4cm}
  \centering
  \subfloat[$\theta_H = 15 \si{\degree}$, $\theta_V = 0 \si{\degree}$, $b = 600
  \si{\degree / \s}$ and $k_v = 0 \si{\N \m / (\radian / \s)}$]{\includegraphics[width=1.\textwidth,
    keepaspectratio]{UPAT_Eye_Model_Passive_Pulleys_v3_State_h15v0kv0.pdf}\label{fig:simulated-saccade-a}} 
  \\
  \subfloat[$\theta_H = 0 \si{\degree}$, $\theta_V = 15 \si{\degree}$, $b = 600
  \si{\degree / \s}$ and $k_v = 0 \si{\N \m / (\radian / \s)}$]{\includegraphics[width=1.\textwidth,
    keepaspectratio]{UPAT_Eye_Model_Passive_Pulleys_v3_Statesh0v15kv0.pdf}\label{fig:simulated-saccade-b}}
  \\
  \subfloat[$\theta_H = -15 \si{\degree}$, $\theta_V = 15 \si{\degree}$, $b = 600
  \si{\degree / \s}$ and $k_v = 0 \si{\N \m / (\radian / \s)}$]{\includegraphics[width=1.\textwidth,
    keepaspectratio]{UPAT_Eye_Model_Passive_Pulleys_v3_Statesh-15v15kv0.pdf}\label{fig:simulated-saccade-c}}
  \\
  \subfloat[$\theta_H = -15 \si{\degree}$, $\theta_V = 15 \si{\degree}$, $b = 100
  \si{\degree / \s}$ and $k_v = 0.002 \si{\N \m / (\radian / \s)}$]{\includegraphics[width=1.\textwidth,
    keepaspectratio]{UPAT_Eye_Model_Passive_Pulleys_v3_Statesh15v-15kv0002.pdf}\label{fig:simulated-saccade-d}}
  \caption{Simulated saccade response for different model parameters. Left
    subplot represents the simulated generalized coordinates, middle the
    generalized velocities and right the estimated \gls{em} excitation
    levels.}\label{fig:simulated-saccade}
\end{figure}

>>>>

\documentclass[11pt,a4paper,draft=false]{report}
\usepackage{amsmath,amsthm,amssymb}
\usepackage{bm,txfonts}       % bm: bold for Greek letters, txfonts -> varmathbb
\usepackage{breqn}              % automatically brake line in equations
\usepackage{cancel}
\usepackage{siunitx}            % include before arydshln
\usepackage{arydshln}         % matrix dash lines
\usepackage{booktabs}         % tables
\usepackage{color,soul}         % used for \hl{}
\usepackage[inline]{enumitem}   % use to set labels in the enumerate environment
\usepackage{fullpage}
\usepackage[pdftex]{graphicx}
\graphicspath{{./figures/}{./tikz/}}
\DeclareGraphicsExtensions{.pdf,.png,.jpg}
\usepackage{wrapfig}
\usepackage{caption}
\captionsetup{labelfont=bf}
\usepackage[caption=false,labelfont=bf]{subfig}
\usepackage{listings}
% \usepackage{algorithm}
% \usepackage{algorithmic}
\usepackage{url}
\usepackage[pdfpagelabels,pdfusetitle,colorlinks=true,pdfborder={0 0
  0},pdfauthor={Dimitar Stanev},colorlinks=false]{hyperref}
\usepackage{natbib}
\bibliographystyle{IEEEtranN}

@@

\author{Konstantinos Filip, Dimitar Stanev\footnote{Electrical and Computer
    Engineering Department, University of Patras, Greece; Corresponding
    author: \url{stanev@ece.upatras.gr}}, and Konstantinos Moustakas}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\begin{abstract}
  Physics-based modeling and dynamic simulation of human eye movements has
  significant implications for improving our understanding of the oculomotor
  system and treating various visuomotor disorders. We introduce an open-source
  biomechanical model of the human eye that can be used for kinematics and
  dynamics analysis. This model is based on the passive pulley hypothesis,
  constructed based on the data reported in literature regarding physiological
  measurements of the human eye and made publicly available\footnote{SimTK
    project: \url{https://simtk.org/projects/eye}}. The model is implemented in
  \texttt{OpenSim}, which is an open-source framework for modeling and
  simulation of musculoskeletal systems. The model incorporates an eye globe,
  orbital suspension tissues and six extraocular muscles. The excitation and
  activation patterns for a variety of targets can be calculated using the
  proposed closed-loop fixation controller that drives the model to perform
  saccadic movements in a forward dynamics manner. The controller minimizes the
  error between the desired saccadic trajectory and the predicted
  movement. Consequently, this model enables the investigation muscle activation
  patterns during static fixation and analyze the dynamics of eye movements.
\end{abstract}

@@

Rapid and accurate eye movements are of great importance for natural vision and
thus studying human eye movement can improve our understanding of the oculomotor
system and treating various visuomotor disorders~\cite{Lee2006, Wei2010a}. Over
the past decades, biomechanics simulation has provided the means to analyze
different human movements~\cite{Delp2007}. The same principles can be used to
analyze visual tasks by modeling the musculoskeletal properties of the
oculomotor system. Consequently, this model can be used to investigate muscle
activation patterns during static fixation, analyze the dynamics of various eye
movements, calculate metabolic costs and simulate eye disorders, such as
different forms of strabismus~\cite{Wong2004}. Furthermore, it can be easily
integrated with available full body models in order to analyze the relation
between the vestibular and oculomotor systems.

Eye movements are a generated from the coordinated activation of the six
\gls{em}. Clinical trials have provided a profound knowledge on the properties
of the \gls{em} and their line of action on the eye globe~\cite{Robinson1969a}
and the resistive tension of the surrounding tissues~\cite{Collins1981,
  Iskander2018}. Various computational models of the extraocular muscles and
orbital mechanics have been proposed, which provide insight for oculomotor
biomechanics, control of eye movement~\cite{Bach-y-Rita1971} and binocular
misalignment. These models focus on the realism of muscle behavior and they were
based on the viscoelastic properties and physiological data \gls{em}.

The first 3D biomechanical model was developed by~\cite{Robinson1964a,
  Robinson1969}, who simplified the formulation by only considering the
elasticity of the \gls{em} ignoring their dynamics. The model incorporates
anatomically realistic muscle paths and empirical innervation-length-tension
relationships. In order to study the neural control of rapid saccadic movements,
models using anatomical and mechanical properties of \gls{em} have been
developed by accounting for the nonlinear muscle dynamics~\cite{Thelen2003a,
  Millard2013}. Such models, having the advantage of supporting dynamics
simulation, are used in conjunction with brain level
controllers~\cite{Angelaki2004, James2018}.

@@

The six \gls{em}, including four rectus muscles and two oblique muscles, are
controlled by the cranial nerves so as to track a visual target and to stabilize
the image of the object of interest on the retina. The \gls{lr} and \gls{mr}
muscles form an agonist/antagonist pair that produce horizontal eye
movements. The \gls{sr} and \gls{ir} muscles form the vertical
agonist/antagonist pair, which mainly controls vertical eye movement and also
affects rotation about the line of sight (secondary action) and the horizontal
plane (tertiary action). The \gls{so} muscle passes through the cartilaginous
trochlea attached to the orbital wall, which reflects the \gls{so} path by
$51 \si{\degree}$. The \gls{io} muscle originates from the orbital wall
anteroinferior to the globe center and inserts on the sclera posterior to the
globe equator. The primary actions of \gls{so} and \gls{io} cause rotation of
the globe around the visual axis, but also affect vertical (secondary action)
and horizontal (tertiary action) movements.

The model relies on the passive pulley assumption, which states that the pulleys
have fixed to the orbit pulley points~\cite{Clark1977,
  Miller2007}. \autoref{tab:eye-muscle-path} shows the positions of the origin,
insertion and pulleys for the \gls{em}, defined in the local body coordinates of
the eye globe. The data are based on physiological
measurements~\cite{Iskander2018}, with some minor modification so as to prevent
unrealistic muscle-surface penetration. Since no position was documented for the
origin of the \gls{so}, a point close to the origins of the rectus muscles was
chosen to match the fiber length in the primary position of the \gls{so} muscle.

\begin{table}[ht]
  \centering
  \caption{Muscle path points for the six \gls{em} defined in the local frame of
    the eye globe (dimensions are given in meters).}\label{tab:eye-muscle-path}
  \resizebox{\textwidth}{!}{
  \begin{tabular}{@{}cccccccccc@{}}
    \toprule
    \textbf{Muscle}
    & \multicolumn{3}{c}{\textbf{Origin}}
    & \multicolumn{3}{c}{\textbf{Pulley}}
    & \multicolumn{3}{c}{\textbf{Insertion}} \\
    \midrule
    & \textit{\textbf{Ox}} & \textit{\textbf{Oy}} & \textit{\textbf{Oz}}
    & \textit{\textbf{Px}} & \textit{\textbf{Py}} & \textit{\textbf{Pz}}
    & \textit{\textbf{Ix}} & \textit{\textbf{Iy}} & \textit{\textbf{Iz}} \\
    \midrule
    \gls{lr} & -0.034 & 0.0006 & -0.013 & -0.0102 & 0.0003 & 0.012 & 0.0065 & 0 & 0.0101 \\
    \gls{mr} & -0.030 & 0.0006 & -0.017 & -0.0053 & 0.00014 & -0.0146 & 0.0088 & 0 & -0.0096 \\
    \gls{sr} & -0.0317 & 0.0036 & -0.016 & -0.0092 & 0.012 & -0.002 & 0.0076 & 0.0104 & 0 \\
    \gls{ir} & -0.0317 & -0.0024 & -0.016 & -0.0042 & -0.0128 & -0.0042 & 0.00805 & -0.0102 & 0 \\
    \gls{so} & 0.0082 & 0.0122 & -0.0152 & -0.030834 & 0.001145 & -0.01644 & 0.0044 & 0.011 & 0.0029 \\
    \gls{io} & 0.0113 & -0.0154 & -0.0111 & -0.00718 & -0.0135 & 0 & -0.008 & 0 & 0.009 \\
    \bottomrule
  \end{tabular}
  }
\end{table}

@@

\begin{table}[ht]
  \centering
  \caption{Millard muscle parameters for the
    \gls{em}.}\label{tab:eye-muscle-parameters}
  \begin{tabular}{@{}cccccccccc@{}}
    \toprule
    \thead{Muscle \\ \quad}
    & \thead{Maximum Isometric \\ Force (N)}  % math \si not in bold
    & \thead{Optimal Fiber \\ Length (m)}
    & \thead{Tendon Slack \\ Length (m)}
    & \thead{Maximum Contraction \\ Velocity (m / s)} \\
    \midrule
    \gls{lr} & 1.4710 & 0.04898 & 0.0084 & 3.8483 \\
    \gls{mr} & 1.5740 & 0.04084 & 0.0038 & 4.6155 \\
    \gls{sr} & 1.1768 & 0.04487 & 0.0054 & 4.2009 \\
    \gls{ir} & 1.4269 & 0.04549 & 0.0048 & 4.1437 \\
    \gls{so} & 0.6031 & 0.03956 & 0.0265 & 4.7648 \\
    \gls{io} & 0.5590 & 0.04110 & 0.0015 & 4.5863 \\
    \bottomrule
  \end{tabular}
\end{table}

The active \gls{fl} and \gls{fpe} characteristic curves of the \gls{em} differ
significantly from those of a skeletal muscle. As shown in
\autoref{fig:millard-curves}, we can fine-tune the curve parameters so as to fit
the experimental data available for the \gls{lr} muscle. The values for the
active \gls{fl} and \gls{fpe} characteristic curves are summarized in
Tables~\ref{tab:eye-fl-curve} and~\ref{tab:eye-fpe-curve}, respectively. We
safely assume that the parameters of the characteristic curves for the other
\gls{em} are the same.

\begin{table}
  \centering
  \parbox{.45\linewidth}{
    \centering
    \caption{Parameters of the active \gls{fl} characteristic curve for the
      \gls{em}.}\label{tab:eye-fl-curve}
    \begin{tabular}{@{}cccccccccc@{}}
      \toprule
      \textbf{Parameter} & \textbf{Value} \\
      \midrule
      min norm active fiber length & 0.55 \\
      transition norm fiver length & 0.7 \\
      max norm active fiver length & 1.8 \\
      shallow ascending slope &  2.4 \\
      minimum value & 0.0 \\
      \bottomrule
    \end{tabular}
  }
  \quad
  \parbox{.45\linewidth}{
    \vspace{-1.5cm}
    \centering
    \caption{Parameters of the \gls{fpe} characteristic curve for the
      \gls{em}.}\label{tab:eye-fpe-curve}
    \begin{tabular}{@{}cccccccccc@{}}
      \toprule
      \textbf{Parameter} & \textbf{Value} \\
      \midrule
      strain at zero force & -0.18 \\
      strain at one norm force & 0.4 \\
      \bottomrule
    \end{tabular}
  }
\end{table}

\begin{figure}[ht]
  \centering
  \subfloat[active \gls{fl} curve]{\includegraphics[width=0.5\textwidth,
    keepaspectratio]{active-force-length-curve.png}\label{fig:active-force-length-curve}}
  \subfloat[\gls{fl} curve]{\includegraphics[width=0.5\textwidth,
    keepaspectratio]{passive-force-length-curve.png}\label{fig:passive-force-length-curve}}
  \caption{The active \gls{fl} and \gls{fpe} curve definition for the Millard
    muscle model as implemented in \texttt{OpenSim}.}\label{fig:millard-curves}
\end{figure}

\gls{em} have a higher fraction of fast twitch fibers and thus different
\gls{fv} behavior, due to different structures compared to skeletal
muscles. Despite that, the default Millard \gls{fv} curve was used for the six
\gls{em}, since the behavior of the selected muscle model depends heavily on the
maximum contraction velocity $v^{\text{max}}$. The maximum muscle contraction
velocity is tuned so as to match the peak velocity of saccadic eye movement
$\omega^{\text{max}} = 15.7 \si{\radian / \s}$ ($900 \si{\degree /
  \s}$). Following this definition, the maximum muscle contraction velocity is
given in optimal fiber length per seconds and it is thus different for each
\gls{em}, as their optimal fiber length is different
($v^{\text{max}} = \omega^{\text{max}} r / l_o^M$). Furthermore, since the optic
nerve is much shorter that the average muscle nerve, activation and deactivation
delays ($\tau_a = \tau_d = 5 \si{\milli\s}$) are smaller. Finally, two
separate wrapping spheres for the rectus muscles and the oblique muscles were
created, to avoid abnormal changes on the \gls{fl} curve as the eye rotates.

@@

The passive connective tissues of the orbit apply a restoring force, which
brings the eye back to the central position when the net force from the \gls{em}
is zero. These tissues include all non-muscular suspensory tissues, such as
Tenon's capsule, the optic nerve, the fat pad and the conjunctiva. The
force-displacement elasticity force can be represented as

\begin{equation}\label{equ:passive-tissue}
  \vec{f}_t = -k_p \vec{q} - k_c \vec{q}^3 - k_d \vec{\dot{q}}
\end{equation}
% 
where, $\vec{f}_t$ is the passive tissue forces,
$k_p= 2.225 \cdot 10^{-3} \si{\N \m / \radian}$,
$k_c= 34.53 \cdot 10^{-3} \si{\N \m / \radian^3}$ and
$k_v= 2 \cdot 10^{-3} \si{\N \m / (\radian / \s)}$ the physiological
constants~\cite{Robinson1969a, Collins1981, Priamikov2016} and
$\vec{q}, \vec{\dot{q}} \inr{3}$ the rotational coordinates and velocities of
the mode. These forces are modeled using \texttt{OpenSim}'s expression based
coordinate force.

@@

A fixation controller that calculates the \gls{em} excitations required to track
a desired saccade was implemented. The controller actuates the model in a
closed-loop \gls{fd} manner. The parameters of the controller are the desired
horizontal and vertical fixation angles, the saccade onset and velocity, and the
gains of \gls{pd} tracking controller. A sigmoid function is used for generating
smooth saccade trajectories in the horizontal and vertical direction, while the
torsional component is maintained close to zero. More formally,

\begin{equation}\label{equ:sigmoid}
  \begin{aligned}
    \theta_d(t) &= \frac{a}{2} \Big(\tanh(b (t - t_0)) + 1\Big) \\
    \dot{\theta}_d(t) &= \frac{a b}{2} \Big(1 - \tanh^2(b (t - t_0))\Big)
  \end{aligned}
\end{equation}
% 
where $\theta_d(t)$ and $\dot{\theta}_d(t)$ represent the desired orientation
and velocity at time $t$, $a$ the magnitude of the trajectory, $b$ the slope and
$t_0$ a time shift constant. Provided a fixation goal $\theta_g$, a desired
saccade velocity $\dot{\theta}_g$ and a saccade onset $t_g$, the parameters of
the sigmoid function are defined as $a = \theta_g$,
$b = 2 \dot{\theta}_g / \theta_g$ and $t_0 = t_s$. The output of the \gls{pd}
tracking controller has the following form

\begin{equation}\label{equ:eye-pd-controller}
  u(t) = k_p (\theta_d(t) - \theta(t)) + k_d (\dot{\theta}_d(t) -
  \dot{\theta}(t))
\end{equation}
% 
where $k_p$, $k_d $ are the tracking gains, and $\theta(t)$, $\dot{\theta}(t)$
the simulated response of the model.

The sign and magnitude of $u(t)$, representing the deviation from
the fixation target for each axis of rotation respectively, are used to
calculate the muscle excitation levels, by assuming that each individual muscle
rotates the eye globe in a particular direction. \autoref{fig:model} presents an
instance of the model during simulation with the corresponding muscles
activated. \autoref{fig:simulated-hor-saccade} depicts the simulated
coordinates, angular velocities and estimated \gls{em} excitation levels that
reproduce the desired saccade trajectory for different model
parameters. Finally, \autoref{fig:simulated-cmplx-saccade} shows alternations in
the saccadic movements both in the horizontal and vertical direction so as to
examine the activation and deactivation patterns of the \gls{em}.

\begin{figure}[ht]
  \centering
  \subfloat{\includegraphics[height=.25\textheight]{model.png}}
  \hspace{3cm}
  \subfloat{\includegraphics[height=.25\textheight]{eye-globe.png}}
  \caption{Model with a fixation target at $\theta_H = -15 \si{\degree}$,
    $\theta_V = 0 \si{\degree}$ during simulation. Blue denotes low and red high
    muscle activation levels.}\label{fig:model}
\end{figure}

\begin{figure}[ht]
  \centering
  \subfloat[$\theta_H = -15 \si{\degree}$, $\theta_V = 0 \si{\degree}$, $b = 600
  \si{\degree / \s}$ and $k_v = 0.002 \si{\N \m / (\radian / \s)}$]
  {\includegraphics[trim={3cm 0cm 2.8cm 0cm},clip,width=1.\textwidth]{horizontal_move.pdf}}  \\
  \subfloat[$\theta_H = -15 \si{\degree}$, $\theta_V = 0 \si{\degree}$, $b = 600
  \si{\degree / \s}$ and $k_v = 0 \si{\N \m / (\radian / \s)}$]
  {\includegraphics[trim={3cm 0cm 2.8cm 0cm},clip,width=1.\textwidth]{abd_no_visc.pdf}} 
  \caption{Simulated saccade response for different model parameters. Left
    subplot represents the simulated generalized coordinates, middle the
    generalized velocities and right the estimated \gls{em} excitation
    levels.}\label{fig:simulated-hor-saccade}
\end{figure}

\begin{figure}[ht]
  \centering
  \includegraphics[trim={3cm 0cm 2.8cm 0cm},clip,width=1.\textwidth]{complex_move.pdf}
  \caption{Simulated saccade response performing abduction/adduction
    ($0-2 \si{\s}$) and supraduction/infraduction ($2-4 \si{\s}$) with
    $k_v = 0.002 \si{\N \m / (\radian /
      \s)}$.}\label{fig:simulated-cmplx-saccade}
\end{figure}

% \begin{figure}[ht]
%   \centering
%   \includegraphics[width=.7\textwidth]{eye-model.png}
%   \caption{Eye model with a fixation $\theta_H = -15 \si{\degree}$,
%     $\theta_V = 15 \si{\degree}$ during simulation. Blue denotes low and red
%     high muscle excitation levels.}\label{fig:model}
% \end{figure}
